IRS-Aided SWIPT: Joint Waveform, Active and Passive Beamforming Design Under Nonlinear Harvester Model

### 2021-08-02 main.tex
:pencil: Update manuscript

		\begin{subsection}{IRS-Aided SWIPT}
			The major issue of WPT and SWIPT is the low harvested DC power, and the passive beamforming provided by the IRS can enhance the end-to-end energy efficiency and boost the harvester input power level. The smart channel control and low power consumption of IRS can bring more opportunities to SWIPT. For multi-user IRS-aided SWIPT systems, dedicated energy beams were proved unnecessary for the Weighted Sum-Power (WSP) maximization problem \cite{Wu2020b} but essential when fairness issue is considered \cite{Tang2019}. It was also claimed that Line-of-Sight (LoS) links could boost the WSP as rank-deficient channels tend to require fewer energy beams \cite{Wu2020a}. However, \cite{Wu2020b,Tang2019,Wu2020a} were based on an inaccurate linear energy harvester model that is known in both the RF and the communication literature to be inefficient and inaccurate \cite{Clerckx2019,Trotter2009,Clerckx2018,Clerckx2016a,Kim2019,Kim2020a,Kim2021,Clerckx2017,Kim2017,Clerckx2018b,Varasteh2020,Varasteh2019d,Varasteh2020a}. In \cite{Xu2021}, the authors proposed a scalable resource allocation framework for SWIPT systems involving large-scale IRS, where the reflectors are grouped into tiles and the optimization process is divided into an offline mode-design stage and an online mode-selection stage to reduce overall complexity. It was concluded that the tile-based two-stage algorithm not only enables a flexible balance between performance and complexity but also adapts well to the physics-based IRS model. A recent tutorial paper \cite{Wu2021c} provided a comprehensive overview on IRS-aided wireless-powered networks, where channel estimation, resource allocation and practical constraints are discussed in detail. To the best of our knowledge, all existing IRS-aided SWIPT papers considered resource allocation and beamforming design for dedicated information and energy users in a single-carrier network. In this paper, we instead build our design based on a proper nonlinear harvester modeling that captures the dependency of the output DC power on both the power and shape of the input waveform, and marry the benefits of joint multi-carrier waveform and active beamforming optimization for SWIPT with the passive beamforming capability of IRS, to investigate the R-E tradeoff for one SWIPT user with practical co-located information decoder and energy harvester. We ask ourselves the important question: \emph{How to jointly exploit the spatial domain and the frequency domain efficiently through joint waveform and beamforming design to enlarge the R-E region of IRS-aided SWIPT?} The contributions of this paper are summarized as follows.

@@

			For the low-complexity design, when the receiver works in PS mode, we instead obtain the phase shift by Algorithm~\ref{al:m_sca} and the waveform amplitude by \eqref{eq:s_i} and \eqref{eq:s_p}. In contrast to the BCD algorithm that obtains the R-E region by varying the rate constraint from maximum capacity $C_{\max}$\footnote{Recall in Remark~\ref{re:subband_tradeoff} that passive beamforming enables a resource allocation opportunity at the channel such that different capacities are achievable.} to \num{0}, the LC-BCD algorithm draws the R-E tradeoff by performing a two-dimensional search that adjusts the combining and splitting ratios from \num{1} to \num{0}. For the WIT point, $C_{\max}$ can be obtained as a special case of LC-BCD algorithm where $\rho=0$ and the objective function becomes $R$ instead of $z$. The steps are summarized in the Low Complexity-BCD (LC-BCD) Algorithm~\ref{al:lc_bcd}. Given a solution accuracy $\epsilon_{\mathrm{IPM}}$ for the interior-point method, the computational complexity of Algorithm~\ref{al:lc_bcd} is $\mathcal{O}\left(I_{\mathrm{LC-BCD}}I_{\mathrm{M-SCA}}(L+1)^{4.5} \log(\epsilon_{\mathrm{IPM}}^{-1})\right)$, where $I_{\mathrm{LC-BCD}}$ denotes the number of LC-BCD iterations \cite{Luo2010}.

@@

@@

		The impacts of the number of transmit antennas $M$ and the IRS elements $L$ on the R-E behavior are revealed in Figs.~\subref*{fi:re_tx} and \subref*{fi:re_reflector}. \emph{First,} adding either active or passive elements enhances the equivalent channel strength and boosts both information and power transfer, but has negligible influence on the optimal receiving strategy. \emph{Second,} the conventional Linear Energy Harvester (LEH) model leads to power-inefficient design. To show the underlying performance loss, we truncate the DC objective function \eqref{eq:z} at $n_0=2$ such that, 1) in the passive beamforming problem, $z(\boldsymbol{\Phi}) = {\beta_2}{\rho}(t_{\mathrm{I},0}+t_{\mathrm{P},0})/2$ and no SCA is required, 2) in the waveform design problem, the WPT-optimal strategy is the Adaptive Single Sinewave (ASS) waveform strategy that allocates all power to the multisine at the strongest subband \cite{Clerckx2016a}. As shown in Figs.~\subref*{fi:scaling_tx} and \subref*{fi:scaling_reflector}, those conventional designs do not exploit the harvester nonlinearity and end up with a nearly \SI{20}{\dB} current gap compared to SMF and GP. \emph{Third,} passive beamforming has a larger array gain and power scaling order than active beamforming. This behavior is more obvious for the case of WIT and WPT. For active beamforming, doubling $M$ brings a \SI{3}{\dB} gain at the output SNR, which corresponds to a transmit array gain of $M$ \cite{Tse2005} and a doubled harvester input power. Thanks to the rectenna nonlinearity, the output DC current ends up with a nearly four-time (\SI{12}{\dB}) increase, which aligns with the active scaling law in the order of $M^2$ \cite{Clerckx2018b}. On the other hand, when the IRS is very close to the AP, doubling $L$ increases the output SNR up to \SI{6}{\dB}, which implies a reflect array gain of $L^2$ \cite{Wu2019}. An interpretation is that the IRS coherently combines the incoming signal with a receive array gain $L$, then performs an equal gain reflection with a transmit array gain $L$. Hence, doubling the number of IRS elements brings a four-fold increase on the received signal power that further amplifies the harvested DC current by \num{16} times (\SI{24}{\dB}), corresponding to a passive scaling law in the order of $L^4$. Compared with active antennas, IRS elements achieve higher array gain and power scaling order, but a very large $L$ is required to compensate the double fading of the auxiliary link. These observations demonstrate the R-E benefit of passive beamforming and emphasize the importance of accounting for the harvester nonlinearity in the passive beamforming design.

>>>>

		\begin{subsection}{IRS-Aided SWIPT}
			The major issue of WPT and SWIPT is the low harvested DC power, and the passive beamforming provided by the IRS can enhance the end-to-end energy efficiency and boost the harvester input power level. The smart channel control and low power consumption of IRS can bring more opportunities to SWIPT. For multi-user IRS-aided SWIPT systems, dedicated energy beams were proved unnecessary for the Weighted Sum-Power (WSP) maximization problem \cite{Wu2020b} but essential when fairness issue is considered \cite{Tang2019}. It was also claimed that Line-of-Sight (LoS) links could boost the WSP as rank-deficient channels tend to require fewer energy beams \cite{Wu2020a}. However, \cite{Wu2020b,Tang2019,Wu2020a} were based on an inaccurate linear energy harvester model that is known in both the RF and the communication literature to be inefficient and inaccurate \cite{Clerckx2019,Trotter2009,Clerckx2018,Clerckx2016a,Kim2019,Kim2020a,Kim2021,Clerckx2017,Kim2017,Clerckx2018b,Varasteh2020,Varasteh2019d,Varasteh2020a}. In \cite{Xu2021}, the authors proposed a scalable resource allocation framework for SWIPT systems involving large-scale IRS, where the reflectors are grouped into tiles and the optimization process is divided into an offline mode-design stage and an online mode-selection stage to reduce overall complexity. It was concluded that the tile-based two-stage algorithm not only enables a flexible balance between performance and complexity but also adapts well to practical IRS and harvester models. A recent tutorial paper \cite{Wu2021c} provided a comprehensive overview on IRS-aided wireless-powered networks, where channel estimation, resource allocation and practical constraints are discussed in detail. To the best of our knowledge, all existing IRS-aided SWIPT papers considered resource allocation and beamforming design for dedicated information and energy users in a single-carrier network. In this paper, we instead build our design based on a proper nonlinear harvester modeling that captures the dependency of the output DC power on both the power and shape of the input waveform, and marry the benefits of joint multi-carrier waveform and active beamforming optimization for SWIPT with the passive beamforming capability of IRS, to investigate the R-E tradeoff for one SWIPT user with practical co-located information decoder and energy harvester. We ask ourselves the important question: \emph{How to jointly exploit the spatial domain and the frequency domain efficiently through joint waveform and beamforming design to enlarge the R-E region of IRS-aided SWIPT?} The contributions of this paper are summarized as follows.

@@

			For the low-complexity design, when the receiver works in PS mode, we instead obtain the phase shift by Algorithm~\ref{al:m_sca} and the waveform amplitude by \eqref{eq:s_i} and \eqref{eq:s_p}. In contrast to the BCD algorithm that obtains the R-E region by varying the rate constraint from maximum capacity $C_{\max}$\footnote{Recall in Remark~\ref{re:subband_tradeoff} that passive beamforming enables a resource allocation opportunity at the channel such that different capacities are achievable.} to \num{0}, the LC-BCD algorithm intends to draw the R-E tradeoff by performing a two-dimensional search that adjusts $\delta$ and $\rho$ from \num{1} to \num{0}. For the WIT point, $C_{\max}$ can be obtained as a special case of LC-BCD algorithm where $\rho=0$ and the objective function becomes $R$ instead of $z$. The steps are summarized in the Low Complexity-BCD (LC-BCD) Algorithm~\ref{al:lc_bcd}. Given a solution accuracy $\epsilon_{\mathrm{IPM}}$ for the interior-point method, the computational complexity of Algorithm~\ref{al:lc_bcd} is $\mathcal{O}\left(I_{\mathrm{LC-BCD}}I_{\mathrm{M-SCA}}(L+1)^{4.5} \log(\epsilon_{\mathrm{IPM}}^{-1})\right)$, where $I_{\mathrm{LC-BCD}}$ denotes the number of LC-BCD iterations \cite{Luo2010}.

@@

		% \begin{figure}[!t]
		% 	\centering
		% 	\subfloat[Channel]{
		% 		\resizebox{0.45\columnwidth}{!}{
		% 			\input{assets/channel_amplitude_100reflectors.tex}
		% 		}
		% 	}
		% 	\subfloat[WIT waveform: modulated]{
		% 		\resizebox{0.45\columnwidth}{!}{
		% 			\input{assets/waveform_wit_modulated_100reflectors.tex}
		% 		}
		% 	}\\
		% 	\subfloat[WPT waveform: modulated]{
		% 		\resizebox{0.45\columnwidth}{!}{
		% 			\input{assets/waveform_wpt_modulated_100reflectors.tex}
		% 		}
		% 	}
		% 	\subfloat[WPT waveform: multisine]{
		% 		\resizebox{0.45\columnwidth}{!}{
		% 			\input{assets/waveform_wpt_multisine_100reflectors.tex}
		% 		}
		% 	}
		% 	\caption{Sorted channel and waveform amplitude for WIT and WPT with and without IRS versus $N$ for $M=1$, $L=100$, $\sigma_n^2=\SI{-40}{\dBm}$, $B=\SI{10}{\MHz}$ and $d_{\mathrm{H}}=d_{\mathrm{V}}=\SI{2}{\meter}$.}
		% 	\label{fi:amplitude}
		% \end{figure}

@@

		The impacts of the number of transmit antennas $M$ and the IRS elements $L$ on the R-E behavior are revealed in Figs.~\subref*{fi:re_tx} and \subref*{fi:re_reflector}. \emph{First,} adding either active or passive elements enhances the equivalent channel strength and boosts both information and power transfer, but has negligible influence on the optimal receiving strategy. \emph{Second,} the conventional Linear Energy Harvester (LEH) model leads to power-inefficient design. To show the underlying performance loss, we truncate the DC objective function \eqref{eq:z} at $n_0=2$ such that, 1) in the passive beamforming problem, $z(\boldsymbol{\Phi}) = {\beta_2}{\rho}(t_{\mathrm{I},0}+t_{\mathrm{P},0})/2$ and no SCA is required, 2) in the waveform design problem, the WPT-optimal strategy is the Adaptive Single Sinewave (ASS) waveform strategy that allocates all power to the multisine at the strongest subband \cite{Clerckx2016a}. As shown in Figs.~\subref*{fi:scaling_tx} and \subref*{fi:scaling_reflector}, those conventional designs do not exploit the harvester nonlinearity and end up with a nearly \SI{20}{\dB} current gap compared to the nonlinear model-based SMF and GP. \emph{Third,} passive beamforming has a larger array gain and power scaling order than active beamforming. This behavior is more obvious for the case of WIT and WPT. For active beamforming, doubling $M$ brings a \SI{3}{\dB} gain at the output SNR, which corresponds to a transmit array gain of $M$ \cite{Tse2005} and a doubled harvester input power. Thanks to the rectenna nonlinearity, the output DC current ends up with a nearly four-time (\SI{12}{\dB}) increase, which aligns with the active scaling law in the order of $M^2$ \cite{Clerckx2018b}. On the other hand, when the IRS is very close to the AP, doubling $L$ increases the output SNR up to \SI{6}{\dB}, which implies a reflect array gain of $L^2$ \cite{Wu2019}. An interpretation is that the IRS coherently combines the incoming signal with a receive array gain $L$, then performs an equal gain reflection with a transmit array gain $L$. Hence, doubling the number of IRS elements brings a four-fold increase on the received signal power that further amplifies the harvested DC current by \num{16} times (\SI{24}{\dB}), corresponding to a passive scaling law in the order of $L^4$. Compared with active antennas, IRS elements achieve higher array gain and power scaling order, but a very large $L$ is required to compensate the double fading of the auxiliary link. These observations demonstrate the R-E benefit of passive beamforming and emphasize the importance of accounting for the harvester nonlinearity in the passive beamforming design.
